---
title: "Credit_Default_Exercise"
author: "Sarah-Katharina Umek"
date: "2021-01-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Consider a credit portfolio with \emph{n} obligors. Given a common (latent) factor $\Psi$ defaults occur independently with probability $Q = p(\Psi )$. Let $D_i$ be the default indicator for obligor \emph{i} ( $D_i= 1$ if obligor \emph{i} defaults, $D_i = 0$ otherwise), and write $M = D_1 + ... + D_n$ for the total number of defaults.\\

Suppose that \emph{Q} has a Beta distribution with parameters \emph{a} and \emph{b}.\\

\begin{itemize}
    \item Take $n=100, a=1$ and $b=9$, and use simulation to determine the default probabilities $\EX (D_i)$, the default correlations $cor(D_i, D_j), i \neq j$, and the distribution of M.
    
   \item Obtain exact expressions for the default probabilities and correlations, and the distribution of M. How well does simulation work?
\end{itemize}

```{r Implementation}
library("ggpubr")

portfolio_simulate <- function(n = 100, a = 1, b = 9){
  
  # Generating a probability modeled as a beta distribution
  psi <- rbeta(1, a, b)
  
  # Generating n Bernoulli trials
  D <- rbinom(n, 1, psi)
  
  return(list(D, sum(D)))
  
}

run_simulation <- function(runs = 1000){
  
  M_val <- rep(0, runs)
  dframe <- data.frame(sample = 1:100)
  for(i in 1:runs){
    res <- portfolio_simulate()
    M_val[i] <- res[[2]]
    dframe[toString(i)] <- res[[1]]
  }
  hist(M_val, breaks = 1000)
  p<- seq(0, 1, length = 1000)
  plot(p, dbeta(p, 1, 9), type = "l")
  
  cor_sum <- 0
  for(i in 1:100){
    for(j in 1:100){
      if(i != j){
        D_x1 <- as.numeric(dframe[i:i,])[2:1001]
        D_x2 <- as.numeric(dframe[j:j,])[2:1001]
        cor_sum <- cor_sum + cor(D_x1, D_x2, method = c("pearson", "kendall", "spearman"))
      }
    }
  }
  # E[D_i] = P(D_i = 1) * 1 + P(D_i = 0) * 0
  #        = P(D_i = 1) * 1
  #        = Sum of all M values / (Total trials)
  #        = sum(M_val) / (1000 * 100)
  Exp_D <- sum(M_val) / (1000 * 100)
  Cor_D <- cor_sum / 9900 
  cat(sprintf("Expected Default: \"%f\" \nCorrelation: \"%f\"\n", Exp_D, Cor_D))
  return(M_val)
  
}

result <- run_simulation()
# Gives me two values: expected default and the other will be the correlation D_i and D_j
```

