---
title: "Fitting_GARCH"
author: "Sarah-Katharina Umek"
date: "2022-10-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 70), tidy = TRUE)
library(readxl)
library(dplyr)
library(knitr)
library(xtable)
library(stargazer)
library(zoo)
library(xts)
library(quantmod)
library(tidyquant)
library(PerformanceAnalytics)
library(knitr)
library(TTR)
library(PortfolioAnalytics)
library(ROI)
library(ROI.plugin.glpk)
library(ROI.plugin.quadprog)
library(ROI.plugin.symphony)
library(fPortfolio)
library(tseries)
library(lmtest)
library(car)
library(forecast)
library(ggplot2)
library(vars)
library(rugarch)
```

Considering a time series observations} $y_t , t = 1, . . . , T$ of your choice and investigate, if predictability in the conditional mean and/or the conditional variance is present:

```{r time series google}

getSymbols("GOOG", from = '2015-10-30', to =  '2022-10-30', warnings= FALSE, auto.assign= TRUE)
google <- Cl(na.omit(GOOG))
google <- na.omit(diff(log(google)))
plot(google, col = "black", main = "Daily GOOG Return", xlab = "Time")
par(mfrow= c(1,2))
acf(google)
acf(google^2)

```

Continuing by estimating an ARMA-GARCH model for this time series and choosing an appropriate model orders.

```{r fit model}
model_specs <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)), mean.model = list(armaOrder = c(1, 1), include.mean = FALSE), distribution.model = "norm")

fit <- ugarchfit(data = google, spec = model_specs)
infocriteria(fit)

model_specs2 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)), mean.model = list(armaOrder = c(0, 0), include.mean = FALSE), distribution.model = "norm")

fit2 <- ugarchfit(data = google, spec = model_specs2)

infocriteria(fit2)

model_specs2_t <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)), mean.model = list(armaOrder = c(0, 0), include.mean = FALSE), distribution.model = "std")

fit2_t <- ugarchfit(data = google, spec = model_specs2_t)

infocriteria(fit2_t)


model_specs3 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)), mean.model = list(armaOrder = c(0, 1), include.mean = FALSE), distribution.model = "norm")

fit3 <- ugarchfit(data = google, spec = model_specs3)
infocriteria(fit3)

model_specs4 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)), mean.model = list(armaOrder = c(1, 0), include.mean = FALSE), distribution.model = "norm")

fit4 <- ugarchfit(data = google, spec = model_specs4)
infocriteria(fit4)

model_specs5 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)), mean.model = list(armaOrder = c(2, 2), include.mean = FALSE), distribution.model = "norm")

fit5 <- ugarchfit(data = google, spec = model_specs5)
infocriteria(fit5)

model_specs6 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(2,2)), mean.model = list(armaOrder = c(1, 1), include.mean = FALSE), distribution.model = "norm")

fit6 <- ugarchfit(data = google, spec = model_specs5)
infocriteria(fit6)

model_specs6_t <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(2,2)), mean.model = list(armaOrder = c(0, 0), include.mean = FALSE), distribution.model = "std")

fit6_t <- ugarchfit(data = google, spec = model_specs2_t)

infocriteria(fit6_t)


plot(fit2_t, which = 10)
plot(fit6_t, which = 10)

```

Using AIC and BIC to compare the models, we end up choosing an ARMA(0,0) GARCH(1,1) (no significant improvement for higher order ARMA models), however there is an improvement sung student-t distributions.  Checking the standardized residuals for the selected model, we see that the ACF looks okay so we proceed with the Thus our model looks as follows:\newline

$$r_t = \mu_t + X_t, \ \ \ \text{with} \  X_t = \sigma_t \varepsilon_t$$
$$\sigma_t^2 = \alpha_0 + \alpha_1 X_{t-1}^2 + \gamma_1 \sigma_{t-1}^2$$

Next we discuss predictability in the light of the estimated parameters for the selected model.

```{r conditional SD}
fit2_t

sd_sample <- sd(google)
sd_sample

sd_estimates <- as.data.frame(tail(sigma(fit2_t), n = 1))[,1]
sd_estimates

plot(fit2_t, which= 3)
abline(h = sd_sample, col = "orange")
abline(h = sd_estimates, col = "darkgreen")
legend("topleft", legend = c("Sample SD", "Estimated SD"), col = c("orange", "darkgreen"), lty = 1:2, cex = 0.8)

plot(sigma(fit2_t))

```

```{r estimating}
forecast_model <- ugarchforecast(fit2_t, data = NULL, n.ahead = 100, n.roll
= 0, out.sample = 100)

plot(forecast_model, which= 1) 
```


