---
title: "Kalman_Filter_Econometrics"
author: "Sarah-Katharina Umek"
date: "2022-11-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 70), tidy = TRUE)
library(readxl)
library(dplyr)
library(knitr)
library(xtable)
library(stargazer)
library(zoo)
library(xts)
library(quantmod)
library(tidyquant)
library(PerformanceAnalytics)
library(knitr)
library(TTR)
library(PortfolioAnalytics)
library(ROI)
library(ROI.plugin.glpk)
library(ROI.plugin.quadprog)
library(ROI.plugin.symphony)
library(fPortfolio)
library(tseries)
library(lmtest)
library(car)
library(forecast)
library(ggplot2)
library(vars)
library(rugarch)
```

Model: 

We have a Gaussian approximation of the distribution of $\varepsilon_t$ involving the first two moments ($\gamma = - \Gamma ' (1)$ is the Euler Mascheroni constant):

$$\mathbb{E}(\varepsilon_t) = \mu_{\varepsilon} = -\gamma - log(2) \approx -1.27036, \ \ \ \ Var(\varepsilon_t) = \frac{\pi^2}{2} $$
yields a representation as a local level model for $h_t$,

$$h_t = h_{t-1} + \eta_t, \ \ \ \eta_t \sim N(0,\theta) $$
$$y_t = h_t + \varepsilon_t , \ \ \ \ N(0, \frac{\pi^2}{2}) $$
involving the transformed outcome varibale $y_t = log(r^2_t) - \mu_{\varepsilon}$.

Using the propagation, prediction and correction equations we can obtain the following Klaman filter:

$$K_t = (P_{t-1|t-1} + \theta) * C^{-1}_{t|t-1} $$

Applying this model to financial returns of our choice to estimate the volatility $h_t$.

$$\mathbb{E}(\sigma_t^2|\sigma_{t-1}^2) = \sigma_{t-1}^2 \mathbb{E}(e_t^{\eta})= \sigma_{t-1}^2 e^{\theta/2} \approx \sigma_{t-1}^2(1+\theta/2)$$

```{r importing data and calculating squared returns}
#using monthly data as that makes the charts more readable.
getSymbols("GOOG", from = '2015-10-30', to =  '2022-10-30', warnings= FALSE, auto.assign= TRUE, periodicity = "monthly")
google <- Cl(na.omit(GOOG))
google <- na.omit(diff(log(google)))
google_sqrt_r <- google^2
```



```{r setting up Kalman filter(1)}
#We need the Euler-Mascheroni constant $\gamma$
gamma <- -digamma(1)

mu_epsilon <- (-gamma - log(2))
mu_epsilon
```

Recall, transformed variable outcome $y_t = log(r_t^2) - \mu_{\varepsilon}$

```{r setting up Kalman filter(2)}
y_t <- log(google_sqrt_r) - mu_epsilon

#setting up the initial distribution
h_0 <- 0 #mean
P_0 <- 1 #variance

#Kalman filter function

kalman_filter <- function(h0, P0, theta, yt){
  #prediction vectors
  h_t <- rep(h0, length(yt))
  P_t <- rep(P0, length(yt))
  #Kalman filter equations:
  for(i in 2:length(yt)){
    #Propagation
    ht_tminus1 <- h_t[i-1]
    Pt_tminus1 <- P_t[i-1] + theta
    #Prediction
    yt_tminus1 <- ht_tminus1
    Ct_tminus1 <- Pt_tminus1 + (pi^2/2)
    #Correction
    Kt <- Pt_tminus1 * (Ct_tminus1)^(-1)
    ht_t <- ht_tminus1 + Kt*(yt[i] - yt_tminus1)
    Pt_t <- (1 - Kt)*Pt_tminus1
    
    #save the calculated values
    h_t[i] <- ht_t
    P_t[i] <- Pt_t
  }
  return(list("ht = " = h_t, "Pt =" = P_t))
}
```

```{r test Kalman filter}

#testing using a theta of 0.01
test_theta_01 <- kalman_filter(h_0, P_0, 0.01, y_t)


sd1_theta_01 <- test_theta_01$`ht = ` + 1.96*sqrt(test_theta_01$`Pt =`)
sd2_theta_01 <- test_theta_01$`ht = ` - 1.96*sqrt(test_theta_01$`Pt =`)

plot(as.numeric(y_t), type = "l", ylab = "observed value", ylim = c(-20,5))
lines(test_theta_01$`ht = `, type = "l", col = "red")
lines(as.numeric(sd1_theta_01), type = "l", lty = 2, col = "blue")
lines(as.numeric(sd2_theta_01), type = "l", lty = 2, col = "blue")

#testing using a theta of 0.1

test_theta_01 <- kalman_filter(h_0, P_0, 0.1, y_t)


sd1_theta_01 <- test_theta_01$`ht = ` + 1.96*sqrt(test_theta_01$`Pt =`)
sd2_theta_01 <- test_theta_01$`ht = ` - 1.96*sqrt(test_theta_01$`Pt =`)

plot(as.numeric(y_t), type = "l", ylab = "observed value", ylim = c(-20,5))
lines(test_theta_01$`ht = `, type = "l", col = "red")
lines(as.numeric(sd1_theta_01), type = "l", lty = 2, col = "blue")
lines(as.numeric(sd2_theta_01), type = "l", lty = 2, col = "blue")


#testing using a theta of 0.5

test_theta_01 <- kalman_filter(h_0, P_0, 0.5, y_t)


sd1_theta_01 <- test_theta_01$`ht = ` + 1.96*sqrt(test_theta_01$`Pt =`)
sd2_theta_01 <- test_theta_01$`ht = ` - 1.96*sqrt(test_theta_01$`Pt =`)

plot(as.numeric(y_t), type = "l", ylab = "observed value", ylim = c(-20,5))
lines(test_theta_01$`ht = `, type = "l", col = "red")
lines(as.numeric(sd1_theta_01), type = "l", lty = 2, col = "blue")
lines(as.numeric(sd2_theta_01), type = "l", lty = 2, col = "blue")

```